---
title: "ilastik tracking of HeLa cells"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  html_notebook:
    code_folding: hide
    source_code: embed
---

Exploration of segmentation, tracking and processing of HeLa cells using Fiji for registration and ilastik for pixel classification, segmentation, tracking and analysis.


```{r "setup", include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8,
                      echo=FALSE, warning=FALSE, message=FALSE,
                      cache=TRUE)
library(tidyverse)

```

```{r get_data}

#Read in a CSV with ilastik tracking data
data_dir <- "/graylab/share/dane/CellTracking/"
nuclear_dir_name <- "Red_Reg"
reporter_dir_name <- "GFP_Reg"
barcode <- ""

beacon_dirs <- dir(paste0(data_dir, barcode), pattern = "Beacon-42", full.names = TRUE)
#Get the nuclear data
raw_data <- lapply(beacon_dirs, function(beacon_dir){
  nuclear_data_filename <- dir(paste0(beacon_dir,"/",nuclear_dir_name),pattern = "CSV-Table", full.names = TRUE)
  if(is_empty(nuclear_data_filename)) stop("No nuclear data found in ", beacon_dir)
  nuclear_data <- read_csv(nuclear_data_filename) 
  nuclear_data$Location <- str_remove(beacon_dir, ".*/")
  nuclear_data$Location_trackId <- paste(nuclear_data$Location, nuclear_data$trackId, sep="_")
  #Get the reporter data for the same image location
  reporter_data_filename <- dir(paste0(beacon_dir,"/",reporter_dir_name),pattern = "CSV-Table", full.names = TRUE)
  if(is_empty(reporter_data_filename)) stop("No reporter data found in ", beacon_dir)
  reporter_data <- read_csv(reporter_data_filename)
  all_data <- nuclear_data %>%
    inner_join(reporter_data, by=c("frame", "Center_of_the_object_0", "Center_of_the_object_1"))
  # missing_nuclear_data <- nuclear_data %>%
  #   anti_join(all_data, by=c("frame", "Center_of_the_object_0", "Center_of_the_object_1"))
  #   missing_reporter_data <- reporter_data %>%
  #   anti_join(all_data, by=c("frame", "Center_of_the_object_0", "Center_of_the_object_1"))
  colnames(all_data) <- str_replace_all(colnames(all_data), "[.]x","_nuclear")
  colnames(all_data) <- str_replace_all(colnames(all_data), "[.]y","_reporter")
  return(all_data)
}) %>%
  bind_rows()

clean_data <- raw_data %>%
  filter(lineageId_nuclear > 1,
         trackId_nuclear > 1) %>% #limit to valid tracks and lineages
  group_by(Location_trackId) %>%
  mutate(Track_length = n()) %>%
  ungroup() %>%
  #remove dividing, disappearing and appearing cells
  filter(Track_length==max(length(unique(frame)))) %>% 
  group_by(Location_trackId) %>%
  mutate(Mean_Intensity_0_nuclear_Sum = sum(Mean_Intensity_0_nuclear),
         Mean_Intensity_0_nuclear_T0 = Mean_Intensity_0_nuclear[frame==0],
         Mean_Intensity_0_nuclear_T0Norm = Mean_Intensity_0_nuclear/Mean_Intensity_0_nuclear_T0,
         Mean_Intensity_0_nuclear_T0Norm_Sum = sum(Mean_Intensity_0_nuclear_T0Norm)) %>%
  mutate(Mean_Intensity_0_reporter_Sum = sum(Mean_Intensity_0_reporter),
         Mean_Intensity_0_reporter_T0 = Mean_Intensity_0_reporter[frame==0],
         Mean_Intensity_0_reporter_T0Norm = Mean_Intensity_0_reporter/Mean_Intensity_0_reporter_T0,
         Mean_Intensity_0_reporter_T0Norm_Sum = sum(Mean_Intensity_0_reporter_T0Norm)) %>%
  ungroup()

```
 
Nuclei motion with a subset of the  data.  
 
```{r samplePlots, fig.width=12, fig.height=6}
 #Debug on a subset of the data
set.seed(42)
clean_data_subset <- sample(unique(clean_data$Location_trackId), size = 1000, replace = FALSE)
clean_data_small <- clean_data %>%
  filter(Location_trackId %in% clean_data_subset) %>%
  arrange(Location, frame)

p <- ggplot(clean_data_small, aes(x = Center_of_the_object_0, y = Center_of_the_object_1, colour = factor(trackId_nuclear))) + 
  geom_path(size=1.2) +
  guides(colour=FALSE) +
  scale_y_reverse() +
  labs(title = "Motion over time") +
  facet_wrap(~Location)
p

```

Cell counts that include divsion, appearance and disappearance. Tracks labeled as invalid in ilastik are removed.  

```{r cellCount, fig.width=8, fig.height=6}

df <- raw_data %>%
  filter(lineageId_nuclear > 1,
         trackId_nuclear > 1) %>%
  group_by(Location, frame) %>%
  summarise(Cell_count = n())

p <- ggplot(df, aes(x = frame, y = Cell_count)) +
  geom_line(alpha=.3, size=.8) +
  coord_cartesian(ylim=c(2000,2300)) +
  guides(colour=FALSE) +
  labs(title = "Cell Count over time",
       y = "Count",
       x = "Frame") +
  facet_wrap(~Location)
p

```

```{r reporterheatmap, fig.width=8, fig.height=12, eval=FALSE}
df <- clean_data

p <- ggplot(df, aes(x = frame, y = log10(Mean_Intensity_0_reporter_T0Norm_Sum), colour = Mean_Intensity_0_reporter_T0Norm)) +
  geom_tile() +
  labs(title = "Normalized reporter intensity",
       y = "Cell track ID",
       x = "Frame") + 
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) +
  facet_wrap(~Location)
p

```


Mean nuclear intensity T0 normalized for all cells that are present during the entire experiment.  
```{r nuclearIntensityByCell, fig.width=8, fig.height=6}
df <- clean_data %>%
  group_by(Location_trackId) %>%
  arrange(Mean_Intensity_0_nuclear_T0Norm_Sum) %>%
  ungroup()

p <- ggplot(df, aes(x = frame, y = log10(Mean_Intensity_0_nuclear_T0Norm), colour = factor(trackId_nuclear))) +
  geom_line(alpha=.3, size=.8) +
    guides(colour=FALSE) +
  labs(title = "nuclear intensity",
       y = "Nuclear intensity (normalized to T0)",
       x = "Frame") + 
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())+
  facet_wrap(~Location)
p <- p + geom_smooth(aes(x = frame,  y = log10(Mean_Intensity_0_nuclear_T0Norm)), colour="black")
p

```

Mean reporter intensity T0 normalized for all cells that are present during the entire experiment.  

```{r reporterIntensityByCell, fig.width=8, fig.height=6}
df <- clean_data %>%
  group_by(Location_trackId) %>%
  arrange(Mean_Intensity_0_reporter_T0Norm_Sum) %>%
  ungroup()

p <- ggplot(df, aes(x = frame, y = log10(Mean_Intensity_0_reporter_T0Norm), colour = factor(trackId_nuclear))) +
  geom_line(alpha=.3, size=.8) +
    guides(colour=FALSE) +
  labs(title = "reporter intensity",
       y = "Reporter intensity (normalized to T0)",
       x = "Frame") + 
  # theme(axis.title.y=element_blank(),
  #       axis.text.y=element_blank(),
  #       axis.ticks.y=element_blank())+
  facet_wrap(~Location)
p <- p + geom_smooth(aes(x = frame,  y = log10(Mean_Intensity_0_reporter_T0Norm)), colour="black")
p

```

